Interactive Visualization for System Log Analysis

### 2013-12-17 paper/paper.tex
Updated related work, began algorithm section

\section{Related Work}
% A description of previous papers related to your project.

\begin{figure*}[p]
    \centering
    \includegraphics[width=0.9\textwidth]{images/ganglia.png}
    \caption{Screenshot of Ganglia\protect\footnotemark}
    \label{fig:ganglia}
\end{figure*}
\footnotetext{\url{http://en.wikipedia.org/wiki/File:ScalableGridEngineGanglia2.png}}
\begin{figure*}[p]
    \centering
    \includegraphics[width=0.9\textwidth]{images/rrdtool.png}
    \caption{Screenshot of RRDtool\protect\footnotemark}
    \label{fig:rrdtool}
\end{figure*}
\footnotetext{\url{http://oss.oetiker.ch/rrdtool/gallery/index.en.html}}

There are many good system monitor solutions out there including Ganglia~\cite{Massie04} and Munin
which is based on RRDtool~\cite{Oetiker99}. However, as you can see in Figure~\ref{fig:ganglia} and
Figure~\ref{fig:rrdtool}, the visualizations become harder to detect anomalies in the system as the
number of nodes increases.

We found Wei Xu and collaborators~\cite{Xu09} made very good system of detecting anomalies by mining
log data. We adopted their machine learning approach including their feature specifications in order
to create an intelligent visualization system.

\section{Methods}
% A detailed explanation of the techniques and algorithms you used to solve the problem.

We have a system which generates such log files using the Google Logging Library
(glog)\footnote{\url{https://code.google.com/p/google-glog/}}. Our goal is to create a generic
visualization library for glog, with emphasis on this system. Theoretically, anybody using glog in a
distributed framework should be able to work with the visualizations out-of-the-box. Here is the
summary of our algorithm:

\paragraph{Algorithm Summary}
\begin{enumerate}
    \item Compute per-node features
        \begin{itemize}
            \item Message count vector
            \item Statistics across time
            \item Each node computes its own features
        \end{itemize}
    \item Select k most representative nodes
        \begin{itemize}
            \item Using distributed k-means clustering
            \item User interactively chooses k
        \end{itemize}
    \item Display interactive visualization
        \begin{itemize}
            \item Client processes visualization
            \item Web server clusters on-demand
            \item Server communicates through Apache server
        \end{itemize}
\end{enumerate}

>>>>

Our system is open source and available on github~\footnote{\url{http://www.github.com/wjkoh/cs262a}}.

\section{Related Work}
Given the wide range of distributed applications, system log analysis has been a prominent research area for decades.
Unfortunately, given the available visualization tools in the past, none are adequate for off-the-shelf use.
We divide the related work into two sections: system state analysis and system log analysis.
The system state refers to aggregating data about the statistics of resource usage in a system, agnostic to the application.
System logs are arbitrary print statements in a piece of code, distributed across one or more nodes.

\subsection{System State Analysis}
Ganglia\footnote{\url{http://www.ganglia.info}} is a system monitoring system which can handle up to 2,000 nodes and aggregate statistics in real-time~\cite{Massie04}.
It requires installing software on each running node and can only handle predefined metrics; it does not allow user-defined metrics and does not utilize system logs.
The data is not clustered either, so either all nodes are visualized, or all nodes are aggregated.
Finally, it is far from interactive: simple options such as setting a time window require many clicks.
See Figure~\ref{fig:ganglia} for a screenshot of the amount of data shown for a single node.

\begin{figure*}[h]
    \centering
    \includegraphics[width=0.9\textwidth]{images/ganglia.png}
    \caption{Screenshot of Ganglia (image courtesy of Wikipedia). This information shows various metrics for a single node.}
    \label{fig:ganglia}
\end{figure*}

\begin{figure*}[p]
    \centering
    \includegraphics[width=0.9\textwidth]{images/rrdtool.png}
    \caption{Screenshot of RRDtool\protect\footnotemark}
    \label{fig:rrdtool}
\end{figure*}
\footnotetext{\url{http://oss.oetiker.ch/rrdtool/gallery/index.en.html}}

Munin, based on RRDtool, is similar to Ganglia: after installation on each node, it automatically presents a visualization of the system status~\cite{Oetiker99}.
As is evident from Figure~\ref{fig:rrdtool}, the visualizations is extremely dense and illegible as the number of nodes increase.

We use ideas from these system state visualizations, and apply them to system logs.
We also recognize that viewing all logs makes an illegible visualization, viewing one node at a time is inefficient, and aggregating all nodes into a single server status hides any anomalous behavior.

\subsection{System Log Analysis}

Wei Xu and collaborators~\cite{Xu09} uses machine learning to detect anomalies by mining log data.
When an anomaly is detected, a decision tree visualization is presented to the user which attempts to explain why it is an anomaly.
This setup, while useful, is fundamentally untrustable: no algorithm can successfully detect all anomalies, and the user will eventually have to resort to command-line \texttt{grep}.
We adopt their machine learning features for our clustering algorithm.

HP Operations Analytics\footnote{\url{http://www8.hp.com/us/en/software-solutions/operations-analytics-operations-analysis/}} provides many similar features to our work, including interactive time-window zooming.
It does not seem to cluster any data, nor does it focus on software bugs, but instead analyze single-node server issues.
It seems to require modifying source code to work with their log analysis system, which may limit its applicabilty.
Unfortunately, it is also closed-source proprietary software so we are not able to test out its effectiveness.

Seaview has the closest goals to ours~\cite{hangalseaview}, and provide similar visualizations.
However, they do not work with arbitrary log types (notably, log messages with multiple variables), nor does it extend to multiple nodes or large amounts of data.
They include no machine learning or reduction of the amount of data shown, other than a planned \texttt{grep} text search.

Makanju and collaborators present spatio-temporal clustering of log data as well \cite{makanju2012spatio}.
Their method does not seem to be, nor claim to be, interactive.
We find their clustering method and features to be complementary to ours, and it would fit cleanly in our interactive system.

\section{Methods}

We automatically extract the source code line which printed each log message, the time it was printed, the node which generated it, and the variable part of each log message.

Our algorithm can be summarized as follows:

\begin{enumerate}
    \item Compute per-node features
    \item Perform distributed k-means clustering
    \item Display interactive visualization
\end{enumerate}

We desribe each of these in detail below.

\subsection{Feature Extraction}
Each node computes its own features, and passes on the feature vector as well as raw log data to the master server.

Following the example of Xu and collaborators \cite{Xu09}, we extract a message count vector in the time window requested by the user.
The message count vector is the number of unique times each message was printed.
This is a strong feature for failing or slowing nodes.
The state ratio vector compares how many times one type of message was printed compared to similar ones; for example, the ratio of file opens to file closes.
We found that the state ratio vector was not useful for our log data, and so we do not compute it to save CPU time.

Using the extracted variables, we compute basic statistics about each type of message.
In a given time window, we compute the:
\begin{enumerate}
\item mean,
\item standard deviation,
\item slope of a linear regression,
\item z-score, and
\item r-value.
\end{enumerate}
These would indicate anomalies involving, for example, a diverging residual in a machine learning application, a section of code which takes too long to run, and a message which keeps working on the same file rather than working on new files.

\subsection{Clustering}
    %\item Perform distributed k-means clustering

\subsection{Visualization}
    %\item Display interactive visualization

%         \begin{itemize}
%             \item Message count vector
%             \item Statistics across time
%             \item Each node computes its own features
%         \end{itemize}
%     \item Select k most representative nodes
%         \begin{itemize}
%             \item Using distributed k-means clustering
%             \item User interactively chooses k
%         \end{itemize}
%     \item Display interactive visualization
%         \begin{itemize}
%             \item Client processes visualization
%             \item Web server clusters on-demand
%             \item Server communicates through Apache server
%         \end{itemize}
%

\section{Implementation}
We identify each log message by the node and source code line that generated it.
We assume that all logs are generated using the Google Logging Library (GLOG)\footnote{\url{https://code.google.com/p/google-glog/}} in order to know which source line generated which log message.
This assumption can be dropped using the mining algorithm of Xu and collaborators~\cite{Xu:2009:OSP:1674659.1677125}.
